The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2024-02-29 analysis/paper/paper.qmd
Update intro text

# Introduction {#intro}

The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft. The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005. The program distinguished itself internationally as a post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field. With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation. All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism. Although located in Europe and named "European", its geographic scope was global, visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.

In 2021, TU Delft stepped out from the program, but its approach has remained imprinted on the TU Delft approach to urbanism. Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and, to some extent, MSc Landscape Architecture tracks. There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit. Understanding what made the program impactful would be beneficial for targeted urbanism education and practice in response to current societal challenges. Having a close look at the assignments that were addressed by the program as well as research and design approaches used to tackle those topics throughout the years can provide valuable insight into what are the core, stable elements of urbanism practice, on the one hand, and how does the field need to adapt to changing societal challenges, on the other.

To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program. To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU: 96 theses with an rounded average of `r round(mean(word_counts, na.rm = TRUE), -3) |> format()` words produced over the years for the duration of the program between `r paste0(min(emu$grad_year, na.rm = TRUE) - 2, "-", max(emu$grad_year, na.rm = TRUE))`. Although the corpus of thesis is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.

@@

**The rise of computational social sciences** [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012]. Often such data is available in unstructured textual form and formats that are not easily machine-readable, such as Adobe PDF or Microsoft Word documents [@benchimol2022]. While quantitative text analysis have been widely adopted in computationla social sciences such as linguistics, communication, and political science [@benchimol2022], urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data. This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories - the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then. Patterns revealed in past education can potentially better inform future decisions in education in the field.

Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021]. Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a]. LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively. Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022a]. Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].

In addition to describing the legacy of the EMU, **this paper showcases a reproducible research workflow**. Together with a published FAIR dataset, it forms a reproducible research compendium. The analyses presented in this paper were carried out in the free and open-source statistical programming software R [@rcoreteam2023]. The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository [@forgaci2023] and it is available for exploration in a Shiny dashboard at <http://mintemu_shiny.bk.tudelft.nl:3838/mintemu/>.

>>>>

## Introduction {#intro}

The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft in 2021. The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005. The program distinguished itself internationally as an advanced post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field. With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation. All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism. Although located in Europe and named "European", its geographic scope was global, visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.

In 2021, TU Delft left the program, but its approach has remained imprinted on the TU Delft approach to urbanism. Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities—mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters—are visible in the current MSc Urbanism and, to some extent, MSc Landscape Architecture tracks. There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit. Understanding what made the program impactful would be beneficial for targeted urbanism education and practice in response to current societal challenges. Having a close look at the assignments that were addressed by the program as well as research and design approaches used to tackle those topics throughout the years can provide valuable insight into what are the core, stable elements of urbanism practice, on the one hand, and how does the field need to adapt to changing societal challenges, on the other.

To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program. To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU program: 96 theses with a rounded average of `r round(mean(word_counts, na.rm = TRUE), -3) |> format(big.mark=",")` words produced over the years for the duration of the program between `r paste0(min(emu$grad_year, na.rm = TRUE) - 2, "-", max(emu$grad_year, na.rm = TRUE))`. Although the corpus of theses is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.

@@

The rise of computational social sciences [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012]. Often such data is available in unstructured textual form and formats that are not easily machine-readable, such as Adobe PDF or Microsoft Word documents [@benchimol2022]. While quantitative text analysis have been widely adopted in computational social sciences such as linguistics, communication, and political science [@benchimol2022], urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data. This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories—the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then. Patterns revealed in past education can potentially better inform future decisions in education in the field.

Acknowledging that opportunity, this paper leverages topic modeling, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021]. Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022]. LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively. Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022]. Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].

In addition to describing the legacy of the EMU, this paper showcases a reproducible research workflow. Together with a published FAIR dataset, it forms a reproducible research compendium. The analyses presented in this paper were carried out in the free and open-source statistical programming software R [@rcoreteam2023]. The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository [@forgaci2023] and it is available for exploration in a Shiny dashboard at <http://mintemu_shiny.bk.tudelft.nl:3838/mintemu/>.
